---
title: "p8105_hw1_nja"
output: github_document
---

Problem 1: create a data frame comprised of:
a random sample of size 8 from a standard Normal distribution
a logical vector indicating whether elements of the sample are greater than 0
a character vector of length 8
a factor vector of length 8, with 3 different factor “levels”

Try to take the mean of each variable in your dataframe. What works and what doesn’t?
```{r}
library(tidyverse)
set.seed(545)
problem1_df = tibble(
  norm_samp = rnorm(8),
  vec_logical = norm_samp >0,
  vec_char = c("The", "dog", "ran", "over", "the", "fence", "right", "now"),
  vec_factor = factor(c("low", "medium", "high", "low", "high", "medium", "low", "medium"))
)
mean(pull(problem1_df, norm_samp))
mean(pull(problem1_df, vec_logical))
mean(pull(problem1_df, vec_char))
mean(pull(problem1_df, vec_factor))

eval = FALSE
as.numeric(pull(problem1_df, vec_logical))
as.numeric(pull(problem1_df, vec_char))
as.numeric(pull(problem1_df, vec_factor))
```
I was able to take the mean of the first 2 variables of norm_samp and vec_logical. I was not able to take the mean of the second 2 variables of vec_char and vec_factor. When I run the as.numeric function on vec_logical, vec_char and vec_factor, I see results for both vec_logical and vec_factor but not vec_char. This is because the function of as.numeric converted both the vec_logical and vec_factor into a coded number. For example, for vec_logical convereted True and False into either 0 or 1, while for vec_factor it converted low, medium and high into either 1, 2 or 3. Since the variable vec_char is a string of random character variables, no nuemeric values are able to be added, which is why the output is NA. This does help explain why we were not able to take the mean above. If we now tried to take a mean of the numeric versions of vev_logical and vec_factor, then we theoretically should be able to take them. 

In a second code chunk:
convert the logical vector to numeric, and multiply the random sample by the result
convert the logical vector to a factor, and multiply the random sample by the result
convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result

```{r}
eval = TRUE
as.numeric(pull(problem1_df, vec_logical))*(pull(problem1_df, norm_samp))
as.factor(pull(problem1_df, vec_logical))*(pull(problem1_df, norm_samp))
as.numeric(as.factor(pull(problem1_df, vec_logical))*(pull(problem1_df, norm_samp)))
```

Problem 2: Create a data frame comprised of:
x: a random sample of size 500 from a standard Normal distribution
y: a random sample of size 500 from a standard Normal distribution
A logical vector indicating whether x + y > 1
A numeric vector created by coercing the above logical vector
A factor vector created by coercing the above logical vector

```{r}

set.seed(1234)
problem2_df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  vec_logical_2 = x + y >1,
  vec_numeric_2 = as.numeric(vec_logical_2),
  vec_factor_2 = as.factor(vec_logical_2)
)
```

Write a short description of your vector using inline R code, including: 
* the size of the dataset (using nrow and ncol) 
* the mean, median, and standard deviation of x 
* the proportion of cases for which x + y > 1

The size of the dataset `problem2_df` is `r nrow(problem2_df)` rows and `r ncol(problem2_df)` columns

The mean of x is `r mean(pull(problem2_df, x))`, 
The median is `r median(pull(problem2_df, x))`, 
The standard deviation is `r sd(pull(problem2_df, x))` 
The proportion of cases for which x + y >1 is `r mean(pull(problem2_df, vec_logical_2))`

Make a scatterplot of y vs x; color points using the logical variable (adding color = ... inside of aes in your ggplot code should help). Make a second and third scatterplot that color points using the numeric and factor variables, respectively, and comment on the color scales

```{r}
ggplot(problem2_df) + geom_point(aes(x = x, y = y), color = 'blue') 
ggsave("scatter_plot.pdf", plot = last_plot())

ggplot(problem2_df) + geom_point(aes(x=vec_numeric_2, y=y), color = 'pink')

ggplot(problem2_df) + geom_point(aes(x=vec_factor_2, y=y), color = 'yellow')
```





